---
title: |
  | IFQuant report
  |
  | \normalsize `r sample.info[sample.info=="sample_id",2]`
author: ""
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry: "left=1.5cm,right=1.5cm,top=2cm,bottom=2cm"
fontsize: 10pt
output:
  pdf_document:
    keep_tex: false
    fig_crop: no
papersize: a4
classoption:
- table
header-includes:
  - \usepackage{graphicx}
  - \usepackage{xcolor}
  - \usepackage{placeins}
  - \usepackage{booktabs}
  - \usepackage{titling}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \pagenumbering{gobble}
  - \setlength{\droptitle}{-1cm}
  - \pretitle{\begin{center}\vspace{5cm}\Huge\sffamily}
  - \posttitle{\end{center}\vspace{1cm}}
  - \preauthor{\begin{center}\sffamily}
  - \postauthor{\end{center}}
  - \predate{\begin{center}\sffamily}
  - \postdate{\end{center}}
  - \renewcommand{\headrulewidth}{0.4pt}
  - \renewcommand{\footrulewidth}{0.4pt}
  - \let\HeadRule\headrule
  - \renewcommand\headrule{\color{gray}\HeadRule}
  - \let\FootRule\footrule
  - \renewcommand\footrule{\color{gray}\FootRule}
  - \fancyhead[L]{\sffamily\footnotesize\textcolor{gray}{`r gsub("([_#$%&])","\\\\\\1",sample.info[sample.info=="sample_id",2])`}}
  - \fancyfoot[L]{\sffamily\footnotesize\textcolor{gray}{`r format(Sys.time(), '%d %b. %Y')`}}
  - \fancyfoot[C]{}
  - \fancyfoot[R]{\sffamily\footnotesize\textcolor{gray}{Page \thepage}}
  - \definecolor{colorsection}{RGB}{22,47,118}
  # style for section 
  - \let\Oldsection\section
  - \renewcommand\section[1]{\Oldsection*{\normalfont\sffamily\huge\textcolor{colorsection}{#1\vspace{3mm}}}\addcontentsline{toc}{section}{#1}} 
  # style for subsection -->
  - \let\Oldsubsection\subsection
  - \renewcommand\subsection[1]{\Oldsubsection{\normalfont\sffamily\LARGE\textcolor{colorsection}{#1}}}
  # style for subsubsection -->
  - \let\Oldsubsubsection\subsubsection
  - \renewcommand\subsubsection[1]{\Oldsubsubsection{\normalfont\sffamily\Large\textcolor{colorsection}{#1}}}
  # style for paragraph -->
  - \let\Oldparagraph\paragraph
  - \renewcommand\paragraph[1]{\Oldparagraph{\sffamily\textcolor{colorsection}{#1}}}
---


\sffamily




```{r,echo=FALSE}
options(knitr.table.format = "latex")
knitr::opts_chunk$set(fig.width=10,fig.height=7,fig.align='left',warning=FALSE) 
```


```{r initialization, include=FALSE}
figdir="figures_tmp/"
##map rgb to colour name
rgb2col=function(r,g,b,maxColorValue = 255){
 color.names=grep("[0-9]",colours(distinct=TRUE),invert=TRUE,value=TRUE)
 color.rgb=t(col2rgb(color.names))/255
 color.names[which.min(sqrt((color.rgb[,"red"]-r/maxColorValue)**2+(color.rgb[,"green"]-g/maxColorValue)**2+(color.rgb[,"blue"]-b/maxColorValue)**2))]
}
```




\newpage
\pagenumbering{arabic}
# Image

```{r image_specifications,echo=FALSE,results='asis'}
tmptable=sample.info
tmptable[,1]=gsub(":*$",":",tmptable[,1])
colnames(tmptable)=NULL
tmptable[,1]=sapply(gsub(" id([ :])"," ID\\1", gsub("_"," ",tmptable[,1])),function(x){substr(x, 1, 1) <- toupper(substr(x, 1, 1));x})
knitr::kable(tmptable,row.names=FALSE, vline = "",toprule="", midrule="",bottomrule="",linesep="")
```

\vfill

Panel specifications: 
\footnotesize
```{r panel_specifications,echo=FALSE,results='asis'}
tmptable=data.frame(Channel=metadata$channel,Filter=metadata$filter,Marker=metadata$name,"Display color"=sapply(strsplit(metadata$color,","),function(x){rgb2col(as.numeric(x[1]),as.numeric(x[2]),as.numeric(x[3]))}),check.names=FALSE,stringsAsFactors=FALSE)
##colorize column Display color
tmptable[,"Display color"]=sapply(tmptable[,"Display color"],function(color){rgb=col2rgb(color);bg="black";if(color=="black")bg="white";paste0("\\cellcolor{",bg,"}\\textcolor[RGB]{",rgb[1,],",",rgb[2,],",",rgb[3,],"}{\\textbf{",color,"}}")})
knitr::kable(tmptable,format="latex",escape=FALSE,row.names=FALSE,booktabs = TRUE,linesep="")
```
\normalsize
\vfill

Composite image (white: selected tiles, red: excluded regions, green: regions of interest`r if(!(clip.region["x.min"]<=0&&clip.region["y.min"]<=0&&clip.region["x.max"]>=image.info$image.width.pixel&&clip.region["y.max"]>=image.info$image.height.pixel)){", grey area: ignored region"}`):
```{r figure_composite, echo = FALSE, message=FALSE, fig.align='left', out.width='100%',out.height='60%'}
knitr::include_graphics(paste0(figdir,"/image_composite.pdf"))
```

<!-- Regions of interest properties -->
```{r ROIs_table,echo = FALSE,results='asis',out.width='100%'}
if(length(regions.of.interest.labels)>0)
{
 cat("\\vspace{3mm}\n")
 cat("Regions of interest:\n")
 cat("\n")
 tmptable=get_formated_ROI_summary_table()
 print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE))
 cat("\n")
}
```

\newpage
# Selected tiles

Specific markers (column) with `r dapi.channel.name``r ifelse(tumor_marker_present,paste0(" and ",ck.channel.name),"")` for selected tiles (row):
```{r figure_selected_tiles, echo = FALSE,results='asis'}
nr=nrow(selected.tiles)
nc=length(rownames(metadata)[!metadata$channel%in%c(channel.dapi,channel.ck,channel.autofluorescence)])
im.w=(0.8/nc)
im.h=(0.9/nr)
tmptable=expand.grid(Region=1:nrow(selected.tiles),f=rownames(metadata)[!metadata$channel%in%c(channel.dapi,channel.ck,channel.autofluorescence)])
tmptable$filename=paste0("\\includegraphics[height=",im.h,"\\textheight,width=",im.w,"\\textwidth]{{",figdir,"/selected_tile_",tmptable$Region,"_",tmptable$f,".pdf}}")
tmptable=reshape2::dcast(tmptable,Region~f,value.var="filename")
##use marker name for colnames
colnames(tmptable)[grep("channel_",colnames(tmptable))]=metadata[colnames(tmptable)[grep("channel_",colnames(tmptable))],"name"]
colnames(tmptable)[1]=""
knitr::kable(tmptable,format="latex",escape=FALSE,row.names=FALSE, vline = "",toprule="",midrule="",bottomrule="",linesep="",valign = "top",align="c") #booktabs = TRUE
```



\newpage
# QC

```{r figure_QC, echo = FALSE, message=FALSE, fig.align='left', out.width='100%',out.height='60%'}
knitr::include_graphics(paste0(figdir,"/image_QC.pdf"))
```

\vspace{1cm}
```{r QC_percent,echo=FALSE,results='asis'}
nb.nonempty.bins=QC.bin.count["ok.sharpness"]+QC.bin.count["warning.sharpness"]+QC.bin.count["bad.sharpness"]
tmptable=data.frame()
tmptable=rbind(tmptable,data.frame(c1="Green", c2=paste0(paste(nucleus.channel.names,collapse="+")," focus OK"),c3=paste0("(",signif(100*QC.bin.count["ok.sharpness"]/nb.nonempty.bins,2),"%)")))
tmptable=rbind(tmptable,data.frame(c1="Yellow",c2=paste0(paste(nucleus.channel.names,collapse="+")," focus warning"),c3=paste0("(",signif(100*QC.bin.count["warning.sharpness"]/nb.nonempty.bins,2),"%)")))
tmptable=rbind(tmptable,data.frame(c1="Red",   c2=paste0(paste(nucleus.channel.names,collapse="+")," out of focus"),c3=paste0("(",signif(100*QC.bin.count["bad.sharpness"]/nb.nonempty.bins,2),"%)")))
tmptable=rbind(tmptable,data.frame(c1="Grey",  c2=paste0("Low ",paste(nucleus.channel.names,collapse="+")),c3=""))
tmptable=rbind(tmptable,data.frame(c1="White", c2="No signal",c3=""))
tmptable=rbind(tmptable,data.frame(c1="Blue",  c2="Excluded regions",c3=""))
colnames(tmptable)=NULL
knitr::kable(tmptable,row.names=FALSE, vline = "",toprule="", midrule="",bottomrule="",	 linesep="")
```





\newpage
# Cell segmentation

`r if(length(nucleus.channel.names)>1){paste0("Markers used for nuclei segmentation: ",paste(nucleus.channel.names,collapse=", "),".")}`

Cell density (cells/mm$^2$):
```{r figure_cell_density, echo = FALSE, message=FALSE, fig.align='left', out.width='100%',out.height='60%'}
knitr::include_graphics(paste0(figdir,"/image_cell_density.pdf"))
```


\newpage
Nuclei centers (green) detected in selected tiles with `r dapi.channel.name``r ifelse(tumor_marker_present,paste0(" and ",ck.channel.name),"")`:
```{r figure_cell_segmentation, echo = FALSE,results='asis'}
n=nrow(selected.tiles)
nc=round(sqrt(n*210/297))
nr=ceiling(n/nc)
im.w=(0.9/nc)
im.h=(0.9/nr)
tmptable=data.frame(i=1:n)
tmptable$filename=paste0("\\includegraphics[height=",im.h,"\\textheight,width=",im.w,"\\textwidth]{{",figdir,"/selected_tile_",tmptable$i,"_cell_segmentation.pdf}}")
tmptable$x=(tmptable$i-1)%%nc
tmptable$y=(tmptable$i-1)%/%nc
tmptable=reshape2::dcast(tmptable,y~x,value.var="filename",fill="")[,-1]
##remove colnames
colnames(tmptable)=NULL
knitr::kable(tmptable,format="latex",escape=FALSE,row.names=FALSE, vline = "",toprule="",midrule="",bottomrule="",linesep="",valign = "top",align="c")
```


\newpage
# Tissue segmentation

```{r tissue_segmentation1,echo = FALSE,results='asis',out.width='100%'}
if(tumor_marker_present==FALSE)
{
  cat("\\textcolor{red}{In the absence of a tumor marker, tissue segmentation could not be done!}")
}
if(tumor_marker_present)
{
  cat("\n")
  cat("Automatic ",ck.channel.name," thresholding:\n",sep="")
  cat("\\begin{minipage}[t]{0.5\\textwidth}\n")
  cat("\\vspace{0pt}\n")

  color="black"
  if(automatic.ck.thresholding.status=="Success")color="green"
  if(automatic.ck.thresholding.status=="Warning")color="orange"
  if(automatic.ck.thresholding.status=="Failed")color="red"
  if(automatic.ck.thresholding.status=="Not done")color="gray"
  if(automatic.ck.thresholding.status=="Manually adjusted")color="gray"
  tmptable=data.frame()
  tmptable=rbind(tmptable,data.frame(c1="Status:",c2=paste0("\\textcolor{",color,"}{",automatic.ck.thresholding.status,"}")))
  tmptable=rbind(tmptable,data.frame(c1=paste0(ck.channel.name," threshold:"),c2=as.character(signif(metadata.channel.thresholding[input.ck,"threshold"],5))))
  colnames(tmptable)=NULL
  print(knitr::kable(tmptable,row.names=FALSE, vline = "",toprule="", midrule="",bottomrule="",linesep="",escape=FALSE))

  cat("\\end{minipage}\n")
  cat("\\begin{minipage}[t]{0.4\\textwidth}\n")
  cat("\\vspace{-1cm}\n")
  cat("\\includegraphics[width=1\\textwidth]{{",figdir,"/automatic_ck_thresholding.pdf}}\n",sep="")
  cat("\\end{minipage}\n")
  cat("\\vspace{5mm}\n")
  tmptable=get_formated_tissue_type_summary_table()
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("Density=(Number of cells in tissue)/(total tissue area)\n")
  cat("\\vspace{5mm}\n")

  ##we cannot set size based on width because one figure has a legend that we want to ignore.
  ##we want to fix the height such that width (ignoring legend) ~ 0.19 \textwidth and height is smaller than 0.25 \textheight
  max.w=0.19 #percent of page width
  max.h=0.25 #percent of page height
  im.h=min(max.h,max.w*clipped.image.height.pixel/clipped.image.width.pixel)

  cat("\\begin{minipage}[t]{0.66\\textwidth}\n")
  cat("Tissue type (red: tumor, green: stroma",ifelse(flag_other_tissue,", blue: other",""),"):\n",sep="")
  cat("\\begin{flushleft}\\includegraphics[width=\\textwidth,height=0.6\\textheight]{{",figdir,"/image_tissue_type.pdf}}\\end{flushleft}\n",sep="")
  cat("\\end{minipage}\n")

  cat("\\begin{minipage}[t]{0.33\\textwidth}\n")
  cat(ck.channel.name," fluorescence (with ",dapi.channel.name,"):\n",sep="")
  cat("\\begin{flushleft}\\includegraphics[width=\\textwidth,height=",im.h,"\\textheight]{{",figdir,"/image_channel_intensity_",input.ck,".pdf}}\\end{flushleft}\n",sep="")
  cat("\\vspace{5mm}\n")
  cat(ck.channel.name,"-positive cells density (cells/mm$^2$):\n",sep="")
  cat("\\begin{flushleft}\\includegraphics[width=\\textwidth,height=",im.h,"\\textheight]{{",figdir,"/image_channel_cell_density_",input.ck,".pdf}}\\end{flushleft}\n",sep="")
  cat("\\end{minipage}\n")
}
```


```{r tissue_segmentation2,echo = FALSE,results='asis',out.width='100%'}
if(tumor_marker_present)
{
  cat("\\newpage\n")
  cat("Tissue segmentation in selected tiles (two images per region):\n\n")
  cat("* ",ck.channel.name," positive (red circles) and ",ck.channel.name," negative cells (white dots) with ",dapi.channel.name," and ",ck.channel.name,".\n",sep="")
  cat("* Tissue type mask (red: tumor, green: stroma",ifelse(flag_other_tissue,", blue: other",""),") with ",dapi.channel.name," and ",ck.channel.name,".\n",sep="")
  n=nrow(selected.tiles)
  nc=round(sqrt(n*210/297))
  nr=ceiling(n/nc)
  im.w=(0.9/nc)/2 ##two image by cell
  im.h=(0.9/nr)
  tmptable=data.frame(i=1:n)
  tmptable$filename=paste0("\\includegraphics[height=",im.h,"\\textheight,width=",im.w,"\\textwidth]{{",figdir,"/selected_tile_",tmptable$i,"_tissue_segmentation_ck.pdf}}\\includegraphics[height=",im.h,"\\textheight,width=",im.w,"\\textwidth]{{",figdir,"/selected_tile_",tmptable$i,"_tissue_segmentation.pdf}}")
  tmptable$x=(tmptable$i-1)%%nc
  tmptable$y=(tmptable$i-1)%/%nc
  tmptable=reshape2::dcast(tmptable,y~x,value.var="filename",fill="")[,-1]
  ##remove colnames
  colnames(tmptable)=NULL
  print(knitr::kable(tmptable,format="latex",escape=FALSE,row.names=FALSE, vline = "",toprule="",midrule="",bottomrule="",linesep="",valign = "top",align="c"))
}
```





\newpage
# Markers thresholding

```{r table_channel_thresholding,echo=FALSE,results='asis'}
tmptable=metadata.channel.thresholding[!metadata.channel.thresholding$channel%in%c(channel.dapi,channel.autofluorescence),]
tmptable$region=gsub("^(circular\\.[^.]*|[^.]*)\\.(nucleus|extended.nucleus|around.nucleus)$","\\2",tmptable$score.type)
tmptable$statistics=gsub("^(circular\\.[^.]*|[^.]*)\\.(nucleus|extended.nucleus|around.nucleus)$","\\1",tmptable$score.type)
##reformat region
tmptable$region=gsub("^nucleus$","nucleus",tmptable$region)
tmptable$region=gsub("^extended.nucleus$","extended nucleus",tmptable$region)
tmptable$region=gsub("^around.nucleus$","around nucleus (doughnut)",tmptable$region)
##reformat statistics
tmptable$statistics=gsub("circular\\.","encircling ",tmptable$statistics)
tmptable$statistics=gsub("q050","median",tmptable$statistics)
tmptable$statistics=gsub("q0*([1-9][0-9]*)","\\1 percentile",tmptable$statistics)
tmptable$statistics=gsub("(^1| 1) ","\\1-st ",tmptable$statistics)
tmptable$statistics=gsub("(^2| 2) ","\\1-nd ",tmptable$statistics)
tmptable$statistics=gsub("([0-9]| [0-9]) ","\\1-th ",tmptable$statistics)
tmptable=tmptable[,c("name","statistics","region","threshold")]
colnames(tmptable)=c("Marker","Summary statistics","Region","Threshold")
knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="")
```

Note: in this section, a cell is considered as \<marker\>-positive if the score of this \<marker\> (summary statistics over the region) is above the threshold, independently of the other markers.
\vspace{1cm}

### Total
\footnotesize
```{r table_markerpositive_summary_total,echo=FALSE,results='asis'}
tmptable=get_formated_single_marker_summary_table(tissue.type="total")
knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r")
```
\normalsize
Density=(number of marker-positive cells)/(total area)
\vspace{5mm}

```{r table_markerpositive_summary_stroma,echo=FALSE,results='asis'}
if(tumor_marker_present)
{
  cat("### Stroma\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_single_marker_summary_table(tissue.type="stroma")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of marker-positive cells)/(total stroma area)\n")
  cat("\\vspace{5mm}\n")
}
```

```{r table_markerpositive_summary_tumor,echo=FALSE,results='asis'}
if(tumor_marker_present)
{
  cat("### Tumor\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_single_marker_summary_table(tissue.type="tumor")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of marker-positive cells)/(total tumor area)\n")
  cat("\\vspace{5mm}\n")
}
```


```{r table_markerpositive_summary_other,echo=FALSE,results='asis'}
if(tumor_marker_present&&flag_other_tissue)
{
  cat("### Other\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_single_marker_summary_table(tissue.type="other")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of marker-positive cells)/(total other area)\n")
}
```



\vspace{1cm}
```{r figure_analysis_image_per_channel, echo = FALSE,results='asis'}
##we want nr row per page. But we cannot set the size based on width because of the legend for the cell density.
##instead we set the height such that the width=0.9/nc * \textwidth
nr=2
nc=2
im.h=(0.9/nc)*clipped.image.height.pixel/clipped.image.width.pixel
im.h2=paste0(im.h,"\\textwidth")
textwidth.approx=210-2*15 #15mm of margin on each side
textheight.approx=297-2*20 #20mm of margin on each side
if(im.h*textwidth.approx>(0.8/nr)*textheight.approx)
{
im.h2=paste0((0.8/nr),"\\textheight")
}
for(f in rownames(metadata)[!metadata$channel%in%c(channel.dapi,channel.autofluorescence)])
{
 tmptable=data.frame(
    paste0("\\includegraphics[height=",im.h2,"]{{",figdir,"/image_channel_intensity_",f,".pdf}}"),     paste0("\\includegraphics[height=",im.h2,"]{{",figdir,"/image_channel_cell_density_",f,".pdf}}"))
 colnames(tmptable)=c(paste0(metadata[f,"name"]," fluorescence (with ",dapi.channel.name,")"),paste0(metadata[f,"name"],"-positive cells density (cells/mm$^2$)"))
 print(knitr::kable(tmptable,format="latex",escape=FALSE,row.names=FALSE, vline = "",toprule="",midrule="",bottomrule="",linesep="",valign = "top",align="c"))
 cat("\n")
}
```


\newpage
For each marker (column), marker-positve cells (red circles) and negative cells (white dots) for selected tiles (rows):
```{r  figure_analysis_selected_tiles_per_channel_positive_cells, echo = FALSE,results='asis'}
nr=nrow(selected.tiles)
nc=length(rownames(metadata)[!metadata$channel%in%c(channel.dapi,channel.autofluorescence)])
im.w=(0.8/nc)
im.h=(0.9/nr)
tmptable=expand.grid(Region=1:nrow(selected.tiles),f=rownames(metadata)[!metadata$channel%in%c(channel.dapi,channel.autofluorescence)])
tmptable$filename=paste0("\\includegraphics[height=",im.h,"\\textheight,width=",im.w,"\\textwidth]{{",figdir,"/selected_tile_",tmptable$Region,"_",tmptable$f,"_positive_cells.pdf}}")
tmptable=reshape2::dcast(tmptable,Region~f,value.var="filename")
##use marker name for colnames
colnames(tmptable)[grep("channel_",colnames(tmptable))]=metadata[colnames(tmptable)[grep("channel_",colnames(tmptable))],"name"]
colnames(tmptable)[1]=""
knitr::kable(tmptable,format="latex",escape=FALSE,row.names=FALSE, vline = "",toprule="",midrule="",bottomrule="",linesep="",valign = "top",align="c")
```



\newpage
# Phenotypes

Phenotypes specification:
```{r table_phenotypes_specifications,echo=FALSE,results='asis'}
tmptable=unique(data.table(phenotypes))
for(col in rownames(metadata)[!metadata$channel%in%c(channel.dapi,channel.autofluorescence)])
{
tmptable=setnames(tmptable[,.(new_col=gsub("^01$","*",paste0(sort(unlist(.SD)),collapse=""))),by=c("label",cols[cols!=col]),.SDcols=col],"new_col",col)
}
tmptable=tmptable[,lapply(.SD,function(x){gsub("^1$","+",gsub("^0$","--",x))}),by=label]
##change colnames to marker names
colnames(tmptable)=ifelse(colnames(tmptable)%in%rownames(metadata),metadata[colnames(tmptable),"name"],colnames(tmptable))
setnames(tmptable,"label","Phenotype")
knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="")
```
Symbols: negative (--), positive(+), negative or positive (*)

`r if(!is.null(overlapping_phenotypes))paste0("Warning: phenotypes are not disjoint (i.e. cells can be be assigned to multiple phenotypes). For example, ",names(overlapping_phenotypes)[[1]]," cells will be assigned to the following phenotypes: ",paste(overlapping_phenotypes[[1]],collapse=", "),".")`


\vspace{2cm}
```{r figure_phenotype_summary_total,echo=FALSE,message=FALSE,out.width='100%'}
knitr::include_graphics(paste0(figdir,"/phenotypes_cell_count.pdf"))
```

\newpage
### Total
\footnotesize
```{r table_phenotype_summary_total,echo=FALSE,results='asis'}
tmptable=get_formated_phenotypes_summary_table(tissue.type="total",phenotypes.type="standard")
knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r")
```
\normalsize
Density=(number of cells with phenotype)/(total area)
\vspace{1cm}


```{r table_phenotype_summary_stroma,echo=FALSE,results='asis'}
if(tumor_marker_present)
{
  cat("### Stroma\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="stroma",phenotypes.type="standard")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of cells with phenotype)/(total stroma area)\n")
  cat("\\vspace{1cm}\n")
}
```

```{r table_phenotype_summary_tumor,echo=FALSE,results='asis'}
if(tumor_marker_present)
{
  cat("### Tumor\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="tumor",phenotypes.type="standard")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of cells with phenotype)/(total tumor area)\n")
  cat("\\vspace{1cm}\n")
}  
```


```{r table_phenotype_summary_other,echo=FALSE,results='asis'}
if(tumor_marker_present&&flag_other_tissue)
{
  cat("### Other\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="other",phenotypes.type="standard")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of cells with phenotype)/(total other area)\n")
  cat("\\vspace{1cm}\n")
}
```

### Total (other phenotypes)
\footnotesize
```{r table_phenotype_summary_total_other,echo=FALSE,results='asis'}
tmptable=get_formated_phenotypes_summary_table(tissue.type="total",phenotypes.type="other")
knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r")
```
\normalsize
Density=(number of cells with phenotype)/(total area)
\vspace{1cm}


```{r table_phenotype_summary_stroma_other,echo=FALSE,results='asis'}
if(tumor_marker_present)
{
  cat("### Stroma (other phenotypes)\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="stroma",phenotypes.type="other")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of cells with phenotype)/(total stroma area)\n")
  cat("\\vspace{1cm}\n")
}
```

```{r table_phenotype_summary_tumor_other,echo=FALSE,results='asis'}
if(tumor_marker_present)
{
  cat("### Tumor (other phenotypes)\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="tumor",phenotypes.type="other")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of cells with phenotype)/(total tumor area)\n")
  cat("\\vspace{1cm}\n")
}  
```

```{r table_phenotype_summary_other_other,echo=FALSE,results='asis'}
if(tumor_marker_present&&flag_other_tissue)
{
  cat("### Other (other phenotypes)\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="other",phenotypes.type="other")
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of cells with phenotype)/(total other area)\n")
  cat("\\vspace{1cm}\n")
}
```

\newpage
Cell density (cells/mm$^2$) per phenotype:

```{r figure_phenotypes_cell_density, echo = FALSE,results='asis'}
##we want nr row per page. But we cannot set the size based on width because of the legend for the cell density.
##instead we set the height such that the widht=0.9/nc * \textwidth
nr=2
nc=2
im.h=(0.8/nc)*clipped.image.height.pixel/clipped.image.width.pixel
im.h2=paste0(im.h,"\\textwidth")
textwidth.approx=210-2*15 #15mm of margin on each side
textheight.approx=297-2*20 #20mm of margin on each side
if(im.h*textwidth.approx>(0.8/nr)*textheight.approx)
{
im.h2=paste0((0.8/nr),"\\textheight")
}
for(j in 1:ceiling(length(phenotypes.list)/2))
{
 i=2*(j-1)+1
 tmptable=data.frame(paste0("\\includegraphics[height=",im.h2,"]{{",figdir,"/image_cell_density_phenotype_",i,".pdf}}"))
 i=2*(j-1)+2
 tmptable=cbind(tmptable,data.frame(
  ifelse(i<=length(phenotypes.list),
   paste0("\\includegraphics[height=",im.h2,"]{{",figdir,"/image_cell_density_phenotype_",i,".pdf}}"),
   "")))
 i=2*(j-1)+1
 colnames(tmptable)[1]=paste0(str_sanitize_latex(phenotypes.list[i])," (cells/mm$^2$)")
 i=2*(j-1)+2
 colnames(tmptable)[2]=""
 if(i<=length(phenotypes.list))
 colnames(tmptable)[2]=paste0(str_sanitize_latex(phenotypes.list[i])," (cells/mm$^2$)")
 print(knitr::kable(tmptable,format="latex",escape=FALSE,row.names=FALSE, vline = "",toprule="",midrule="",bottomrule="",linesep="",valign = "top",align="c"))
 cat("\n")
}

```




<!-- ROIs -->
```{r ROIs,echo = FALSE,results='asis',out.width='100%'}
roi.label.idx.list=c(seq_along(regions.of.interest.labels))
if(length(regions.of.interest.labels)>1)roi.label.idx.list=c(roi.label.idx.list,list(seq_along(regions.of.interest.labels)))
if(tumor_tissue_only.ROI.tokeep%in%regions.of.interest.labels)roi.label.idx.list=c(roi.label.idx.list,-1)
for(roi.label.idx in roi.label.idx.list)
{

 cat("\\newpage\n")

 if(length(roi.label.idx)==1&&roi.label.idx>0)
 cat("# Region of interest",roi.label.idx,paste0("(",regions.of.interest.labels[roi.label.idx],")"),"\n")

 if(length(roi.label.idx)==1&&roi.label.idx==-1)
 {
   if(any(tumor_tissue_only.ROI.toremove%in%regions.of.interest.labels))
   {
    cat("# Region of interest",tumor_tissue_only.ROI.tokeep," only ",paste0("(",tumor_tissue_only.ROI.tokeep," without ",paste(tumor_tissue_only.ROI.toremove[tumor_tissue_only.ROI.toremove%in%regions.of.interest.labels],collapse=", "),")"),"\n")
    cat("Region of interest Tumor tissue (green) without",paste(tumor_tissue_only.ROI.toremove[tumor_tissue_only.ROI.toremove%in%regions.of.interest.labels],"(red)",collapse=", "),"\n")
   }
   else
   {
    cat("# Region of interest",tumor_tissue_only.ROI.tokeep," only ",paste0("(",tumor_tissue_only.ROI.tokeep,")"),"\n")
   }
 }	 

 if(length(roi.label.idx)>1)
 cat("# Regions of interest",paste(roi.label.idx,collapse=" & "),paste0("(",paste(regions.of.interest.labels[roi.label.idx],collapse=" & "),")"),"\n")

 cat("\n")
 cat("\\includegraphics[width=\\textwidth,height=0.5\\textheight]{{",figdir,"/image_composite_roi_",paste(roi.label.idx,collapse="_"),".pdf}}\n",sep="")
 cat("\n")

if(tumor_marker_present)
{
 cat("## Tissue segmentation\n")

 cat("\n")		
 tmptable=get_formated_tissue_type_summary_table(roi.label.idx=roi.label.idx)
 print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
 cat("\n")
 cat("Density=(Number of cells in tissue)/(total tissue area)\n")
 cat("\n")
}

 cat("## Markers thresholding\n")

 cat("### Total\n")
 cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_single_marker_summary_table(tissue.type="total",roi.label.idx=roi.label.idx)
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("Density=(number of marker-positve cells)/(total area)\n")
  cat("\\vspace{1cm}\n")
  cat("\\normalsize\n")
  cat("\n")

  if(tumor_marker_present)
  {
    cat("### Stroma\n")
    cat("\\footnotesize\n")
    cat("\n")
    tmptable=get_formated_single_marker_summary_table(tissue.type="stroma",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("Density=(number of marker-positve cells)/(total stroma area)\n")
    cat("\\vspace{1cm}\n")
    cat("\\normalsize\n")
    cat("\n")

    cat("### Tumor\n")
    cat("\\footnotesize\n")
    cat("\n")
    tmptable=get_formated_single_marker_summary_table(tissue.type="tumor",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("Density=(number of marker-positve cells)/(total tumor area)\n")
    cat("\\vspace{1cm}\n")
    cat("\\normalsize\n")
    cat("\n")

   if(flag_other_tissue)
   {
    cat("### Other\n")
    cat("\\footnotesize\n")
    cat("\n")
    tmptable=get_formated_single_marker_summary_table(tissue.type="other",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("Density=(number of marker-positve cells)/(total other area)\n")
    cat("\\vspace{1cm}\n")
    cat("\\normalsize\n")
    cat("\n")
   }
  }
  
 cat("## Phenotypes\n")
 cat("\\vspace{1cm}\n")
 
 cat("\n")
 cat("\\includegraphics[width=\\textwidth]{{",figdir,"/phenotypes_cell_count_roi_",paste(roi.label.idx,collapse="_"),".pdf}}\n",sep="")
 cat("\n")

 cat("### Total\n")

 cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="total",phenotypes.type="standard",roi.label.idx=roi.label.idx)
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("Density=(number of cells with phenotype)/(total area)\n")
  cat("\\vspace{1cm}\n")
  cat("\\normalsize\n")
  cat("\n")

  if(tumor_marker_present)
  {
    cat("### Stroma\n")
    cat("\\footnotesize\n")
    cat("\n")
    tmptable=get_formated_phenotypes_summary_table(tissue.type="stroma",phenotypes.type="standard",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("\\normalsize\n")
    cat("Density=(number of cells with phenotype)/(total stroma area)\n")
    cat("\\vspace{1cm}\n")
    cat("\n")
    
    cat("### Tumor\n")
    cat("\\footnotesize\n")
    cat("\n")
     tmptable=get_formated_phenotypes_summary_table(tissue.type="tumor",phenotypes.type="standard",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("\\normalsize\n")
    cat("Density=(number of cells with phenotype)/(total tumor area)\n")
    cat("\\vspace{1cm}\n")
    cat("\n")
   
    if(flag_other_tissue)
    {
    cat("### Other\n")
    cat("\\footnotesize\n")
    cat("\n")
     tmptable=get_formated_phenotypes_summary_table(tissue.type="other",phenotypes.type="standard",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("\\normalsize\n")
    cat("Density=(number of cells with phenotype)/(total other area)\n")
    cat("\\vspace{1cm}\n")
    cat("\n")
    }
  }

  cat("### Total (other phenotypes)\n")
  cat("\\footnotesize\n")
  cat("\n")
  tmptable=get_formated_phenotypes_summary_table(tissue.type="total",phenotypes.type="other",roi.label.idx=roi.label.idx)
  print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
  cat("\n")
  cat("\\normalsize\n")
  cat("Density=(number of cells with phenotype)/(total area)\n")
  cat("\\vspace{1cm}\n")
  cat("\n")

 if(tumor_marker_present)
 {
    cat("### Stroma (other phenotypes)\n")
    cat("\\footnotesize\n")
    cat("\n")
    tmptable=get_formated_phenotypes_summary_table(tissue.type="stroma",phenotypes.type="other",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("\\normalsize\n")
    cat("Density=(number of cells with phenotype)/(total stroma area)\n")
    cat("\\vspace{1cm}\n")
    cat("\n")
    
    cat("### Tumor (other phenotypes)\n")
    cat("\\footnotesize\n")
    cat("\n")
    tmptable=get_formated_phenotypes_summary_table(tissue.type="tumor",phenotypes.type="other",roi.label.idx=roi.label.idx)
    print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
    cat("\n")
    cat("\\normalsize\n")
    cat("Density=(number of cells with phenotype)/(total tumor area)\n")
    cat("\\vspace{1cm}\n")
    cat("\n")

    if(flag_other_tissue)
    {
     cat("### Other (other phenotypes)\n")
     cat("\\footnotesize\n")
     cat("\n")
     tmptable=get_formated_phenotypes_summary_table(tissue.type="other",phenotypes.type="other",roi.label.idx=roi.label.idx)
     print(knitr::kable(tmptable,row.names=FALSE,booktabs = TRUE,linesep="",escape=FALSE,align="r"))
     cat("\n")
     cat("\\normalsize\n")
     cat("Density=(number of cells with phenotype)/(total other area)\n")
     cat("\\vspace{1cm}\n")
     cat("\n")
    }
  }
}
```
